The Hot Hand Fallacy in Call of Duty Search and Destroy

### 2022-08-30 paper/paper.Rmd
final? with gt

---
title: "The Hot Hand Fallacy in Call of Duty Search and Destroy"
link-citations: true
authors:
  - name: Person One
    email: personone@domain.com
  - name: Person Two
    email: persontwo@domain.com
abstract: |
  Our research investigates patterns in round win percentages in professional Search and Destroy (SnD) matches of the popular first-person shooter game Call of Duty (CoD). First, we find evidence supporting the hypothesis that round win probability can be modeled as a constant across rounds in the series, although not at the naive 50%. Second, we examine the proportion of round wins immediately following a win streak. Given streak length, series length, and series winner, we find no evidence that post-streak win proportion is significantly different than expected, suggesting that the perception of momentum at the team level is deceiving. Wald-Wolfowitz run tests also fail to provide evidence for the hot hand phenomenon.
bibliography: references.bib
keywords:
  - esports
  - Call of Duty
  - hot hand
  - runs test
output:
  rticles::arxiv_article:
    includes:
      in_header: preamble.tex
---

@@

[^1]: SnD bears resemblance to "Bomb Defusal" in Counter-Strike and "Plant/Defuse" in Valorant, two other FPS games played in more popular professional leagues.

@@

-   Each round has a 90 second time limit, not counting the potential extension granted if a bomb is planted
-   A bomb plant takes five seconds. The timer resets if the player stops planting site prior to completing it.
-   A bomb defuse takes seven seconds. The timer resets if the player "drops" the bomb.
-   Once the bomb is planted, the round timer is set to 45 seconds.

@@

We adopt the terminology "series" to refer to a single best-of-11 matchup, so as to mirror the terminology of playoff series in professional leagues like the National Basketball Association (NBA), National Hockey League (NHL), and Major League Baseball (MLB). A "game" or a "match" in such leagues is analogous to a "round" of CoD SnD for the purposes of this paper. Note that SnD is not the only game mode played in competitive CoD.[^4] Teams play in a head-to-head, best-of-five format, where SnD is always played as the second and fifth game modes. The best-of-five matchup could also be called a "series", but since we analyze only the SnD games, we refer to the SnD games as series.

@@

[^7]: Raw data: <https://linktr.ee/CDLArchive>. Cleaned data: <https://github.com/%7Bauthor%7D/%7Brepo%7D/blob/master/data/%7Bfile%7D.csv>.

@@

[^9]: Offensive round win percentage has been nearly constant across the three games during the CDL era: 47.2% in Modern Warfare; 47.9% in Cold War; and 48.1% in Vanguard

@@

This format has minimized the range of rates at which teams start series playing offense. The highest rate at which a team has started their series playing defense across all tournament and qualifier SnD series in a single season is 66% (31 of 47). The lowest is 38% (16 of 42).

@@

Other research has borrowed techniques from the field of quality control, such as identifying unlikely sequences of events with the Wald-Wolfowitz runs test (Peel and Clauset 2015, Steeger et al. 2021). Peel and Clauset found no evidence for unlikely sequences of scoring events in the NHL, College Football, and National Football League, although did in the NBA. As a check on their entropy approach to momentum identification, Steeger et al. found several NHL teams with sequences of wins in the 2018-2019 regular season that violated the Wald-Wolfowitz null hypothesis.

@@

[^12]: Professional CoD has a set of maps (at least three, depending on the season) on which teams choose to play SnD. Most are asymmetric, introducing slight advantages for the offense or defense that do not exist on other maps. Counter-Strike and Valorant also have their own sets of maps for their analogous SnD modes.

@@

As implied by MS's Theorem 1, we should expect the proportion of round wins immediately following a streak of $k$ rounds wins for a series ending in $r$ rounds, $\hat{P}^{+|k,r}_{MS}$, to be strictly less than the observed proportion of round wins given the series length, but not given knowledge of the potential existence of a streak, $P^{+|r}$, i.e. the analogue of the expected proportion in Equation \ref{eq:pwr} for the observed data. However, our context is fundamentally different from that of MS, who evaluate data in controlled settings, unaffected by opposition.[^13]

@@

Let us consider another form of the expected proportion of rounds won immediately after a streak of $k$ round wins in a best-of-$s$ series ending in $r$ rounds: the "notional" proportion

```{=tex}
\begin{equation}\label{eq:pwkr}
\hat{P}_0(\text{win} | k, r) = \hat{P}^{+|k,r}_0 := \begin{cases}
\frac{m - k}{r - k} & \text{team wins series}, r > m; \\
1 & \text{team wins series}, r = m; \\
\frac{s - m - k}{r - k} & \text{team loses series}, r > m + k; \\
0 & \text{team loses series}, r = m + k,
\end{cases}
\end{equation}
```
where $r \in R$ and $m$ from Equation \ref{eq:m}.

@@

We can further decompose Equation \ref{eq:pwkr} by the round $i$ (where $k < i \leq r$) in which the streak of length $k$ carries into.

```{=tex}
\begin{equation}\label{eq:pwkri}
\hat{P}_0(\text{win} | k, r, i) = \hat{P}^{+|k,r,i}_0 := \begin{cases}
\frac{m-k-1}{r-k-1} & i < r-1; \\
\frac{1}{2} & i = r-1; \\
1 & \text{team wins series}, i = r; \\
0 & \text{team loses series}, i = r. \\
\end{cases}
\end{equation}
```
Again, we can apply a binomial test to evaluate the hypothesis that the expected proportion, $\hat{P}^{+|k,r,i}_0$, is equal to the observed proportion, $P^{+|k,r,i}$.

@@

We can attempt to detect the hot hand phenomenon more broadly by focusing on non-random sequences identified by the Wald-Wolfowitz runs test. Note that the probability of success and failure need not be equal; thus, we can incorporate our findings regarding constant round win probability, specifying that the probability of success is $\phi_2 = 0.575$.

Under the null hypothesis for the Wald-Wolfowitz runs test, the number of runs in a sequence of $r$ trials is a random variable that can take on values $+$ or $-$ and arrive at $r^+$ successes (and $r^-$ failures), with the following mean $\mu_r$ [^14] and variance $\sigma_r^2$:

[^14]: We calculate $\mu$ and $\sigma^2$ for each number of rounds $r \in R$.

@@

We can subset the observed series sequences, $\zeta_j$, to those that violate the null hypothesis for the runs test and perform a test of equal proportions, where the null hypothesis is that the observed proportion of a sequence relative to all possible sequences, $P^{\zeta}_j$ ( $\frac{N^\zeta_j}{\sum_j N^\zeta }$), is equal to the expected proportion of the sequence relative to all possible sequences, $\hat{P}^\zeta_j$. The test statistic is

```{=tex}
\begin{equation}\label{eq:prop}
z^\zeta_j = \frac{P^\zeta_j - \hat{P}^\zeta_j}{\sqrt{P^\zeta_{j,\delta} (1 - P^\zeta_{j,\delta}) (1 / N^\zeta_j + 1 / \hat{N}^\zeta_j)  } }
\end{equation}
```
where $N^\zeta_j$ and $\hat{N}^\zeta_j$ respectively represent the observed and expected number of occurrences of sequence $\zeta_j$, and where $P^\zeta_{j,\delta} = (P^\zeta_j - \hat{P}^\zeta_j) / ({N^\zeta_j - \hat{N}^\zeta_j})$. The null hypothesis is that the test statistic is normally distributed. If we can reject the null hypothesis for such sequences, we can build an argument in support of the hot hand effect.[^15]

[^15]: As with the binomial tests for the notional and MS post-streak proportions, we adjust the $p$-values with the BY correction.

@@

Table \ref{tbl:prob-series-lasts-r-rounds} shows the expected proportion of series ending in $r$ rounds given $\phi_0 = 0.5$, $\hat{\Phi}_0(r)$, as well as the observed proportions, $\Phi(r)$.

@@

\caption{Alternate estimates of the constant round win probability ($\phi$) and their corresponding $\chi^2$ values, applying the three methods suggested by Mosteller to CoD SnD. Additionally, the naive constant round win probability $\phi_0$ and its $\chi^2$.}

\centering
\begin{tabular}{lrr}
\toprule
Method & $\phi$ & $\chi^2$ ($p$-value) \\
\midrule

$\phi_0$: Naive & 0.5000 & 16.0 ($\leq$ 0.01) \\
$\phi_1$: Method of moments & 0.5725 & 3.6 (0.6) \\
$\phi_2$: Maximum likelihood & 0.5750 & 3.5 (0.62) \\
$\phi_3$: Minimum $\chi^2$ & 0.5775 & 3.5 (0.62) \\

@@

\end{table}
```
Table \ref{tbl:alternative-constant-ps} shows the expected proportions of series ending in $r$ rounds, $\hat{\Phi}_2(r)$, when re-applying Equation \ref{eq:series-length} for the maximum likelihood estimate $\phi_2 = 0.575$.[^16] We observe that $\hat{\Phi}_2(r)$ more closely matches $\Phi(r)$ than $\hat{\Phi}_0(r)$.

[^16]: The method of moments and minimum $\chi^2$ estimates for $\phi$ are omitted simply because the results would be nearly identical to those for the maximum likelihood estimate of $\phi$ (since they are all $\approx 0.575$).

@@

\centering
\begin{tabular}{rrrr}
\toprule
$r$ & $\Phi(r)$ & $\hat{\Phi}_0(r)$ = 0.5 & $\hat{\Phi}_2(r)$ = 0.575 \\
\midrule

@@

cod_round_win_prop_after_k_wins |> filter(win_round, notional_p_value_adj < 0.05)
cod_round_win_prop_after_k_wins |> filter(win_round, ms_p_value_adj < 0.05)
```

Given that people typically perceive streaks as beginning after the third success (or failure) at minimum [@carlson2007], we focus on streaks of three round wins. Table \ref{tbl:pwkr} compares the observed round win proportion, $P^{+|k=3,r}$, with the notional proportion, $\hat{P}^{+|k=3,r}$, following streaks of $k=3$ round wins. The MS proportion, $\hat{P}^{+|k=3,r}_{MS}$, is also shown as a reference.

@@

\caption{For streaks of $k=3$ round wins in CoD SnD series ending in $r$ rounds, the observed count of rounds wins, $r^{+|k=3,r}$, and proportion of round wins, $P^{+|k=3,r}$, among $N^{k=3,r}$ chances ($P^{+|k=3,r} = r^{+|k=3,r} / N^{k=3,r}$). Additionally, the notional and MS expected proportions, $\hat{P}^{+|k=3,r}_0$ and $\hat{P}^{+|k=3,r}_{MS}$, respectively.
}

\centering
\begin{tabular}{rcrrrrr}
\toprule
$r$ & \text{Win series?} & $r^{+|k=3,r}$ & $N^{k=3,r}$ & $P^{+|k=3,r}$ & $\hat{P}^{+|k=3,r}_0$ & $\hat{P}^{+|=k=3,r}_{MS}$ \\ 
\midrule

7 & yes & 156 & 209 & 74.6\% & 75.0\% & 75.7\% \\ 
8 & yes & 130 & 209 & 62.2\% & 60.0\% & 61.9\% \\ 
9 & yes & 100 & 193 & 51.8\% & 50.0\% & 52.7\% \\ 
10 & no & 8 & 60 & 13.3\% & 14.3\% & 26.7\% \\ 
10 & yes & 66 & 151 & 43.7\% & 42.9\% & 44.7\% \\ 
11 & no & 31 & 129 & 24.0\% & 25.0\% & 30.8\% \\ 
11 & yes & 60 & 150 & 40.0\% & 37.5\% & 38.8\% \\

@@

\end{table}
```
All binomial test $p$-values for both $\hat{P}^{+|k=3,r}_0$ and $\hat{P}^{+|k=3,r}_{MS}$ are found to be insignificant at the $\alpha = 0.05$ threshold, implying that we cannot reject the null hypothesis that the notional and MS post-streak round win proportions are different than the observed proportion for streaks of three.[^17] When performing the same tests for streaks of two, four, and five, there is no case in which we can reject the null hypothesis for $\hat{P}^{+|k,r}_0$. The null hypothesis can only be rejected in some extreme cases for $\hat{P}^{+|k,r}_{MS}$.[^18]

[^17]: $p$-values are not shown for readability purposes.

[^18]: These cases are: (1) $k = 4$, $r = 7$ ($N^{k,r} = 108$); (2) $k = 4$, $r = 10$ ($N^{k,r} = 45$); (3) $k = 5, r = 7$ ($N^{k,r} = 37$); (4) $k = 5, r = 8$ ($N^{k,r} = 4$). In each case, the team won the series.

@@

cod_round_win_prop_after_k_wins_i_rounds |> 
  filter(notional_p_value < 0.05)
cod_round_win_prop_after_k_wins_i_rounds |> 
  filter(notional_p_value_adj < 0.05)
```

Table \ref{tbl:pw3ri} shows the observed and notional proportions of round wins---$P^{+|k=3,r,i}$ and $\hat{P}^{+|k=3,r,i}_0$ respectively---immediately following streaks of three round wins when explicitly accounting for the post-streak round index, $i$.[^19][^20] While some of the observed and notional proportions differ strongly, the BY-corrected $p$-values indicate that the null hypothesis cannot be rejected for any combination of $i$ and $r$ (including those not shown).[^21] The same is true for streaks of two, four, and five.

[^19]: We omit the MS proportions $\hat{P}^{+|k=3,r,i}_{MS}$ since sample sizes are small and results may or may not be valid due to the questionable i.i.d assumption.

[^20]: Some combinations of $r$ and $i$ are not shown in the table for brevity.

[^21]: There are combinations for which we could reject the null hypothesis if we did apply the BY correction, such as for $r = 10$ and $i = 9$.

@@

\caption{For streaks of $k=3$ round wins in CoD SnD series ending in $r$ rounds, the observed count of rounds wins, $r^{+|k=3,r,i}$, and proportion of round wins, $P^{+|k=3,r,i}$, among $N^{k=3,r,i}$ chances, given the index of the round, $i$, immediately following the streak. Additionally, the notional proportion, $\hat{P}^{+|k=3,r,i}_0$. Table restricted to $r \in [10, 11]$ and $r - 3 \leq i \leq r - 1$.}

\centering
\begin{tabular}{rrcrrrr}
\toprule
$r$ & $i$ & \text{Win series?} & $r^{+|k=3,r,i}$ & $N^{k=3,r,i}$ & $P^{+|k=3,r,i}$ & $\hat{P}^{+|k=3,r,i}_0$\\ 
\midrule

10 & 7 & no & 2 & 6 & 33.3\% & 33.3\% \\ 
10 & 7 & yes & 11 & 27 & 40.7\% & 33.3\% \\ 
10 & 8 & no & 1 & 13 & 7.7\% & 33.3\% \\ 
10 & 8 & yes & 9 & 24 & 37.5\% & 33.3\% \\ 
10 & 9 & no & 2 & 12 & 16.7\% & 50.0\% \\ 
10 & 9 & yes & 4 & 20 & 20.0\% & 50.0\% \\ 
11 & 8 & no & 2 & 17 & 11.8\% & 28.6\% \\ 
11 & 8 & yes & 3 & 19 & 15.8\% & 28.6\% \\ 
11 & 9 & no & 5 & 12 & 41.7\% & 28.6\% \\ 
11 & 9 & yes & 10 & 20 & 50.0\% & 28.6\% \\ 
11 & 10 & no & 5 & 14 & 35.7\% & 50.0\% \\ 
11 & 10 & yes & 5 & 22 & 22.7\% & 50.0\% \\

@@

cod_round_arrangements_mle |> 
  filter(!is.na(p_value)) |> 
  filter(is_significant) |> 
  mutate(
    total_actual = sum(n_actual),
    total_expected_adj = sum(n_expected_adj)
  ) |>
  mutate(
    prop_p_value = vectorized_prop_test(
      n_actual, 
      total_actual, 
      round(n_expected_adj), 
      round(total_expected_adj)
    )$p.value,
    orig_prop_p_value = prop_p_value,
    across(prop_p_value, ~p.adjust(.x, method = 'BY') |> round(3))
  )
```

>>>>

---
title: "The Hot Hand Fallacy in Call of Duty Search and Destroy"
link-citations: true
authors:
  - name: Person One
    email: personone@domain.com
  - name: Person Two
    email: persontwo@domain.com
abstract: |
  Our research investigates patterns in round win percentages in professional Search and Destroy (SnD) matches of the popular first-person shooter game Call of Duty (CoD). First, we find evidence supporting the hypothesis that round win probability can be modeled as a constant across rounds in the series, although not at the naive 50%. Second, we examine the proportion of round wins immediately following a win streak. Given the round index, streak length, series length, and series winner, we find no evidence that post-streak win proportion is significantly different than expected, suggesting that the perception of momentum at the team level is deceiving. Finally, we find that sequences of round wins that are identified as non-random by the Wald-Wolfowitz run tests do not occur more or less frequently than expectated, further failing to provide evidence for the hot hand effect.
bibliography: references.bib
keywords:
  - esports
  - Call of Duty
  - hot hand
  - runs test
output:
  rticles::arxiv_article:
    includes:
      in_header: preamble.tex
---

@@

[^1]: SnD bears resemblance to "Bomb Defusal" in Counter-Strike and "Plant/Defuse" in Valorant, two other FPS games played in popular professional leagues.

@@

    -   Each round has a 90 second time limit, not counting the potential extension granted if a bomb is planted
    -   A bomb plant takes five seconds. The timer resets if the player stops planting site prior to completing it.
    -   A bomb defuse takes seven seconds. The timer resets if the player "drops" the bomb.
    -   Once the bomb is planted, the round timer is set to 45 seconds, regardless of the prior round timer.

@@

We adopt the terminology "series" to refer to a single best-of-11 matchup, so as to mirror the terminology of playoff series in professional leagues like the National Basketball Association (NBA), National Hockey League (NHL), and Major League Baseball (MLB). A "game" or a "match" in such leagues is analogous to a "round" of CoD SnD for the purposes of this paper.

Note that SnD is not the only game mode played in competitive CoD.[^4] Teams play in a head-to-head, best-of-five format, where SnD is always played as the second and fifth game modes. The best-of-five matchup could also be called a "series", but since we analyze only the SnD games, we refer to the SnD games as series.

@@

[^7]: Raw data: <https://linktr.ee/CDLArchive>. Cleaned data: <https://github.com/author/repo/blob/master/data/file.csv>.

@@

[^9]: Offensive round win percentage has been nearly constant across the three games during the CDL era: 47.2% in the 2020 season; 47.9% in the 2021 season; and 48.1% in the 2022 season.

@@

    This format has minimized the range of rates at which teams start series playing offense. The highest rate at which a team has started their series playing defense across all tournament and qualifier SnD series in a single season is 66% (31 of 47 series). The lowest is 38% (16 of 42 series).

@@

Other research has borrowed techniques from the field of quality control, such as identifying unlikely sequences of events with the Wald-Wolfowitz runs test. Peel and Clauset [-@peel2015] found no evidence for unlikely sequences of scoring events in the NHL, College Football, and National Football League, although did in the NBA. As a check on their entropy approach to momentum identification, Steeger et al. [-@steeger2021] found several NHL teams with sequences of wins in the 2018-2019 regular season that violated the Wald-Wolfowitz null hypothesis.

@@

[^12]: Professional CoD has a set of maps (at least four, depending on the season) on which teams choose to play SnD. Most are asymmetric, introducing slight advantages for the offense or defense that do not exist on other maps. Counter-Strike and Valorant also have their own sets of maps for their analogous SnD modes.

@@

As implied by MS's Theorem 1, we should expect the proportion of round wins immediately following a streak of $k$ rounds wins for a series ending in $r$ rounds, $\hat{P}^{+|k,r}_{MS}$, to be strictly less than the observed proportion of round wins given the series length, where knowledge of the potential existence of a streak is not given, $P^{+|r}$, i.e. the analogue of the expected proportion in Equation \ref{eq:pwr} for the observed data. However, our context is fundamentally different from that of MS, who evaluate data in controlled settings, unaffected by opposition.[^13]

@@

Let us consider another form of the expected proportion of rounds won immediately after a streak of $k$ round wins in a best-of-$s$ series ending in $r$ rounds, the "notional" proportion. We define the notional proportion as

```{=tex}
\begin{equation}\label{eq:pwkr}
\hat{P}_0(\text{win} | k, r) = \hat{P}^{+|k,r}_0 := \begin{cases}
1 & \text{team wins series}, r = m; \\
0 & \text{team loses series}, r = m + k; \\
\frac{m - k}{r - k} & \text{team wins series}, r > m; \\
\frac{s - m - k}{r - k} & \text{team loses series}, r > m + k,
\end{cases}
\end{equation}
```
where $r \in R$ and $m$ from Equation \ref{eq:m}.[^14]

[^14]: Note that some cases are invalid, such as $\text{team loses series}$, $r < m + k$. It is impossible for team to go on a streak of $k$ round wins and then lose the series in less than $m + k$ rounds.

@@

We can further decompose Equation \ref{eq:pwkr} by the round $i$ ( $k < i \leq r$) in which the streak of length $k$ carries into:

```{=tex}
\begin{equation}\label{eq:pwkri}
\hat{P}_0(\text{win} | k, r, i) = \hat{P}^{+|k,r,i}_0 := \begin{cases}
1 & \text{team wins series}, i = r \text{ or } r = m; \\
0 & \text{team loses series}, i = r \text{ or } r = m + k; \\
\frac{m-k-1}{r-k-1} & i < r-1; \\
\frac{1}{2} & i = r-1.
\end{cases}
\end{equation}
```
where $r \in R$ and $m$ from Equation \ref{eq:m}.[^15][^16] Again, we can apply a binomial test to evaluate the hypothesis that the expected proportion, $\hat{P}^{+|k,r,i}_0$, is equal to the observed proportion, $P^{+|k,r,i}$.

[^15]: The case where $i = r - 1$ is the round prior to the known last round in the series. We know the result of the last round since we know which team won the series. The penultimate round is essentially treated as a toss-up, hence $\frac{1}{2}$.

[^16]: As with Equation \ref{eq:pwkr}, there are cases that are impossible, such as $\text{team loses series}$, $r < m + k$.

@@

We can attempt to detect the hot hand phenomenon more broadly by focusing on non-random sequences identified by the Wald-Wolfowitz runs test. Note that the probability of success and failure need not be equal; thus, we can incorporate our findings regarding constant round win probability.

Under the null hypothesis for the Wald-Wolfowitz runs test, the number of runs in a sequence of $r$ trials is a random variable that can take on values $+$ or $-$ and arrive at $r^+$ successes (and $r^-$ failures), with the following mean $\mu_r$ [^17] and variance $\sigma_r^2$:

[^17]: We calculate $\mu$ and $\sigma^2$ for each number of rounds $r \in R$.

@@

We can subset the observed series sequences, $\zeta_j$, to those that violate the null hypothesis for the runs test and perform a test of equal proportions, where the null hypothesis is that the observed proportion of a sequence relative to all possible sequences, $P^{\zeta}_j$ ( $N^\zeta_j / \sum_j N^\zeta_j$), is equal to the expected proportion of the sequence relative to all possible sequences, $\hat{P}^\zeta_j$. The test statistic is

```{=tex}
\begin{equation}\label{eq:prop}
z^\zeta_j = \frac{P^\zeta_j - \hat{P}^\zeta_j}{\sqrt{P^\zeta_{j,\delta} (1 - P^\zeta_{j,\delta}) (1 / N^\zeta_j + 1 / \hat{N}^\zeta_j)  } }
\end{equation}
```
where $N^\zeta_j$ and $\hat{N}^\zeta_j$ respectively represent the observed and expected number of occurrences of a sequence $\zeta_j$, and where $P^\zeta_{j,\delta} = (P^\zeta_j - \hat{P}^\zeta_j) / ({N^\zeta_j - \hat{N}^\zeta_j})$. The null hypothesis is that the test statistic is normally distributed. If we can reject the null hypothesis for such sequences, we can build an argument in support of the hot hand effect.[^18]

[^18]: As with the binomial tests for the post-streak proportions, we adjust the $p$-values with the BY correction.

@@

Table \ref{tbl:prob-series-lasts-r-rounds} shows the expected proportion of series ending in $r$ rounds given $\phi_0$, $\hat{\Phi}_0(r)$, as well as the observed proportions, $\Phi(r)$.

@@

\caption{Alternate estimates of the constant round win probability ($\phi$) and their corresponding $\chi^2$ values, applying the three methods suggested by Mosteller to CoD SnD. Additionally, the naive constant round win probability $\phi_0$ and its $\chi^2$. $chi^2$ test $p$-values shown in parentheses.}

\centering
\begin{tabular}{lrr}
\toprule
Method & $\phi$ & $\chi^2$ \\
\midrule

$\phi_0$: Naive & 0.5000 & 16.0 ($\leq$ 0.01) \\
$\phi_1$: Method of moments & 0.5725 & 3.6 (0.60) \\
$\phi_2$: Maximum likelihood & 0.5750 & 3.5 (0.62) \\
$\phi_3$: Minimum $\chi^2$ & 0.5775 & 3.5 (0.62) \\

@@

\end{table}
```
Table \ref{tbl:alternative-constant-ps} shows the expected proportions of series ending in $r$ rounds, $\hat{\Phi}_2(r)$, when re-applying Equation \ref{eq:series-length} for the maximum likelihood estimate $\phi_2 = 0.575$.[^19] We observe that $\hat{\Phi}_2(r)$ more closely matches $\Phi(r)$ than $\hat{\Phi}_0(r)$.

[^19]: The method of moments and minimum $\chi^2$ estimates for $\phi$ are omitted simply because the results would be nearly identical to those for the maximum likelihood estimate of $\phi$ (since they are all $\approx 0.575$).

@@

\centering
\begin{tabular}{rrrr}
\toprule
$r$ & $\Phi(r)$ & $\hat{\Phi}_0(r)$ & $\hat{\Phi}_2(r)$ \\
\midrule

@@

cod_round_win_prop_after_k_wins |> filter(win_round, notional_p_value_adj < 0.05)
cod_round_win_prop_after_k_wins |> filter(win_round, ms_p_value_adj < 0.05)

cod_round_win_prop_after_k_wins |> 
  filter(win_streak == 3L) |> 
  mutate(
    across(
      matches('p_value'), 
      ~case_when(
        .x <= 0.01 ~ '<=0.01', 
        TRUE ~ sprintf('%.2f', .x)
      )
    ),
    prop_lab = sprintf('%.1f', 100 * prop),
    notional_lab = sprintf('%.1f (%s)', 100 * notional_prop, notional_p_value_adj),
    ms_lab = sprintf('%.1f (%s)', 100 * ms_prop, ms_p_value_adj)
  ) |>
  arrange(n_rounds, win_series, win_round) |> 
  filter(win_round) |> 
  transmute(
    n_rounds,
    across(win_series, ~ifelse(.x, 'yes', 'no')),
    n,
    total,
    prop_lab,
    notional_lab,
    ms_lab
  ) |> 
  gt::gt() |> 
  gt::cols_label(
    n_rounds = 'r',
    win_series = 'Win series?',
    n = 'r+',
    total = 'N',
    prop_lab = 'O',
    notional_lab = 'a',
    ms_lab = 'b'
  ) |> 
  gt::as_latex() |> 
  cat()
```

Given that people typically perceive streaks as beginning after the third success (or failure) at minimum [@carlson2007], we focus on streaks of three round wins. Table \ref{tbl:pwkr} compares the observed round win proportion, $P^{+|k=3,r}$, with the notional proportion, $\hat{P}^{+|k=3,r}_0$, following streaks of $k=3$ round wins. The MS proportion, $\hat{P}^{+|k=3,r}_{MS}$, is also shown as a reference.

@@

\caption{For streaks of $k=3$ round wins in CoD SnD series ending in $r$ rounds, the observed count of rounds wins, $r^{+|k=3,r}$, and proportion of round wins, $P^{+|k=3,r}$, among $N^{k=3,r}$ chances ($P^{+|k=3,r} = r^{+|k=3,r} / N^{k=3,r}$). Additionally, the notional and MS expected proportions, $\hat{P}^{+|k=3,r}_0$ and $\hat{P}^{+|k=3,r}_{MS}$, respectively. BY-corrected binomial test $p$-values shown in parentheses.}

\centering
\begin{tabular}{rcrrrrr}
\toprule
$r$ & \text{Win series?} & $r^{+|k=3,r}$ & $N^{k=3,r}$ & $P^{+|k=3,r}$ \% & $\hat{P}^{+|k=3,r}_0$ \% & $\hat{P}^{+|=k=3,r}_{MS}$ \% \\ 
\midrule

7 & yes & 156 & 209 & 74.6 & 75.0 (1.00) & 75.7 (1.00) \\ 
8 & yes & 130 & 209 & 62.2 & 60.0 (1.00) & 61.9 (1.00) \\ 
9 & yes & 100 & 193 & 51.8 & 50.0 (1.00) & 52.7 (1.00) \\ 
10 & no & 8 & 60 & 13.3 & 14.3 (1.00) & 26.7 (0.54) \\ 
10 & yes & 66 & 151 & 43.7 & 42.9 (1.00) & 44.7 (1.00) \\ 
11 & no & 31 & 129 & 24.0 & 25.0 (1.00) & 30.8 (1.00) \\ 
11 & yes & 60 & 150 & 40.0 & 37.5 (1.00) & 38.8 (1.00) \\

@@

\end{table}
```
All binomial test $p$-values for both $\hat{P}^{+|k=3,r}_0$ and $\hat{P}^{+|k=3,r}_{MS}$ are found to be insignificant at the $\alpha = 0.05$ threshold, implying that we cannot reject the null hypothesis that the notional and MS post-streak round win proportions are different than the observed proportion for streaks of three. When performing the same tests for streaks of two, four, and five, there is no case in which we can reject the null hypothesis for $\hat{P}^{+|k,r}_0$. The null hypothesis can only be rejected in some extreme cases for $\hat{P}^{+|k,r}_{MS}$.[^20]

[^20]: These cases are: (1) $k = 4$, $r = 7$ ($N^{k,r} = 108$); (2) $k = 4$, $r = 10$ ($N^{k,r} = 45$); (3) $k = 5, r = 7$ ($N^{k,r} = 37$); (4) $k = 5, r = 8$ ($N^{k,r} = 4$). In each case, the team won the series.

@@

cod_round_win_prop_after_k_wins_i_rounds |> filter(notional_p_value < 0.05)
cod_round_win_prop_after_k_wins_i_rounds |> filter(notional_p_value_adj < 0.05)

cod_round_win_prop_after_k_wins_i_rounds |> 
  filter(
    n_rounds >= 7,
    (round + 2) >= n_rounds,
    win_streak == 3L
  ) |> 
  mutate(
    across(
      matches('p_value'), 
      ~case_when(
        .x <= 0.01 ~ '<=0.01', 
        TRUE ~ sprintf('%.2f', .x)
      )
    ),
    prop_lab = sprintf('%.1f', 100 * prop),
    notional_lab = sprintf('%.1f (%s)', 100 * notional_prop, notional_p_value_adj)
  ) |>
  arrange(n_rounds, win_series, win_round) |> 
  filter(win_round) |> 
  transmute(
    n_rounds,
    round,
    across(win_series, ~ifelse(.x, 'yes', 'no')),
    n,
    total,
    prop_lab,
    notional_lab
  ) |> 
  arrange(n_rounds, round, win_series) |> 
  gt::gt() |> 
  gt::cols_label(
    n_rounds = 'r',
    round = 'i',
    win_series = 'Win series?',
    n = 'r+',
    total = 'N',
    prop_lab = 'O',
    notional_lab = 'a'
  ) |> 
  gt::as_latex() |> 
  cat()
```

Table \ref{tbl:pw3ri} shows the observed and notional proportions of round wins---$P^{+|k=3,r,i}$ and $\hat{P}^{+|k=3,r,i}_0$ respectively---immediately following streaks of three round wins when explicitly accounting for the post-streak round index, $i$.[^21] While some of the observed and notional proportions differ strongly, the BY-corrected $p$-values indicate that the null hypothesis cannot be rejected for any combination of $i$ and $r$, including those not shown for brevity.[^22] The same is true for streaks of two, four, and five.

[^21]: We omit the MS proportions $\hat{P}^{+|k=3,r,i}_{MS}$ since sample sizes are small and results may or may not be valid due to the questionable i.i.d assumption.

[^22]: There are combinations for which we could reject the null hypothesis if we did not apply the BY correction, such as for $r = 10$ and $i = 9$.

@@

\caption{For streaks of $k=3$ round wins in CoD SnD series ending in $r$ rounds, the observed count of rounds wins, $r^{+|k=3,r,i}$, and proportion of round wins, $P^{+|k=3,r,i}$, among $N^{k=3,r,i}$ chances, given the index of the round, $i$, immediately following the streak. Additionally, the notional proportion, $\hat{P}^{+|k=3,r,i}_0$. Table restricted to $r \in [7, 8, 9, 10, 11]$ and $r - 3 \leq i \leq r - 1$. BY-corrected $p$-values shown in parentheses.}

\centering
\begin{tabular}{rrcrrrr}
\toprule
$r$ & $i$ & \text{Win series?} & $r^{+|k=3,r,i}$ & $N^{k=3,r,i}$ & $P^{+|k=3,r,i}$ \% & $\hat{P}^{+|k=3,r,i}_0$ \% \\ 
\midrule

7 & 5 & yes & 36 & 54 & 66.7 & 66.7 (1.00) \\ 
7 & 6 & yes & 34 & 54 & 63.0 & 50.0 (1.00) \\ 
8 & 6 & yes & 20 & 41 & 48.8 & 50.0 (1.00) \\ 
8 & 7 & yes & 24 & 46 & 52.2 & 50.0 (1.00) \\ 
9 & 7 & yes & 15 & 33 & 45.5 & 40.0 (1.00) \\ 
9 & 8 & yes & 18 & 37 & 48.6 & 50.0 (1.00) \\ 
10 & 8 & no & 1 & 13 & 7.7 & 33.3 (1.00) \\ 
10 & 8 & yes & 9 & 24 & 37.5 & 33.3 (1.00) \\ 
10 & 9 & no & 2 & 12 & 16.7 & 50.0 (1.00) \\ 
10 & 9 & yes & 4 & 20 & 20.0 & 50.0 (1.00) \\ 
11 & 9 & no & 5 & 12 & 41.7 & 28.6 (1.00) \\ 
11 & 9 & yes & 10 & 20 & 50.0 & 28.6 (1.00) \\ 
11 & 10 & no & 5 & 14 & 35.7 & 50.0 (1.00) \\ 
11 & 10 & yes & 5 & 22 & 22.7 & 50.0 (1.00) \\

@@

cod_round_arrangements_mle |> 
  filter(!is.na(p_value)) |> 
  filter(is_significant) |> 
  mutate(
    total_actual = sum(n_actual),
    total_expected_adj = sum(n_expected_adj)
  ) |>
  mutate(
    prop_p_value = vectorized_prop_test(
      n_actual, 
      total_actual, 
      round(n_expected_adj), 
      round(total_expected_adj)
    )$p.value,
    across(prop_p_value, list(adj = ~p.adjust(.x, method = 'BY'))),
    across(
      matches('p_value'), 
      ~case_when(
        .x <= 0.01 ~ '<=0.01', 
        TRUE ~ sprintf('%.2f', .x)
      )
    ),
    across(ws, ~str_replace_all(.x, '-', ' ') |> str_replace_all(c('1' = '+', '0' = '-'))),
    actual_lab = sprintf('%.2f', 100 * prop_actual),
    expected_lab = sprintf('%.2f (%s)', 100 * prop_expected, prop_p_value)
  ) |> 
  select(
    n_rounds, 
    ws,
    n_actual,
    actual_lab,
    n_expected,
    expected_lab
  ) |> 
  arrange(n_rounds, ws) |> 
  filter(n_rounds <= 10) |> 
  gt::gt() |> 
  gt::cols_label(
    n_rounds = 'r',
    ws = 'zeta',
    n_actual = 'n1',
    actual_lab = 'p1',
    n_expected = 'n2',
    expected_lab = 'p2'
  ) |> 
  gt::as_latex() |> 
  cat()
```
